### 2017-04-24 Conferences/CoSSURF2017/abstract.tex
A skeleton for the CoSSURF abstract.

%\documentclass[11pt]{article}
\documentclass[12pt,letterpaper,oneside,openright]{book}
\usepackage[letterpaper]{geometry}
\usepackage[none]{hyphenat}
\usepackage{libertine}

\begin{document}
\chapter*{\centering Abstract}
\begin{sloppypar}
The NOvA (Neutrino Oscillation ) experiment was set out to measure the neutrino mass hierarchy, the neutrino CP-violating phase $\delta_{CP}$, and the precision measurement of the octant of the mixing angle $\theta_{23}$. NOvA utilizes the Fermilab NuMI beam, and places one near detector at Fermilab and one far detector at Minnisota, 800 miles away from Fermilab. NOvA's most recent results show a non-maximal mixing angle $\theta_{23}$. 
\end{sloppypar}
\end{document}